---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
 
# njmunicipalities

<!-- badges: start -->
[![R-CMD-check](https://github.com/tor-gu/njmunicipalities/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/tor-gu/njmunicipalities/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

This is a data package for R that contains every county
and municipality in New Jersey, from 2000 to 2021.

## Installation

You can install the development version of njmunicipalities from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("tor-gu/njmunicipalities")
```

## Changes in NJ municipalites, 2000-2021

Over the period 2000-2021, there have been several changes to
the list of municipalities in New Jersey:

* In 2005, South Belmar became Lake Como. The US Census assigned
a new GEOID to Lake Como.
* In 2007, Dover township in Ocean County became Toms River and was assigned a new GEOID.
* In 2008, Washington township in Mercer County became Robbinsville township and was assigned a new GEOID.
* In 2009, West Paterson became Woodland Park, and was assigned a new GEOID.
* In 2010, Caldwell borough was assigned a new GEOID from the the US Census, though there was no name change at this time.
* In 2013, Princeton borough and Princeton township merged. The merged municipality retained the Princeton borough GEOID, though the US Census started using the name 'Princeton' in place of 'Princeton borough' for the merged municipality.

This package will return municipality tables for any year from 2000 to
2021, and provides tools for dealing the changes from year to year.

## Examples
#### Get a table of municipalities
Function `get_municipalities` returns a table of municipalities for a
given year. The default is year 2021.
```{r}
library(njmunicipalities)

# Municipality table for 2021
get_municipalities() |> head(n=5)
```
If the year is specified (from 2000 to 2021), the table will reflect the names and GEOIDs in effect for that year. Here is the list for 2007.
```{r}
# Municipality list for 2007
get_municipalities(2007) |> head(n=5)
```

#### Get a table of municipalities for one year with GEOIDs from another
If the optional parameter `geoid_year` is included, the GEOIDs in the returned table will be the ones in effect for that year.  For example, to get a municipality list with municipal names effective 2002 and GEOIDs effective 2018:
```{r}
# Municipality list for 2002 with GEOIDs from 2018
get_municipalities(2002, geoid_year = 2018) |> head(n=5)
```

If you need both GEOIDs, specify `geoid_ref_as_ref_column = TRUE`. This will cause the GEOIDs from `geoid_year` to be returned as a separate column (instead of replacing the `GEOID`).
```{r}
# Municipality list for 2002 with GEOIDs from 2018 added as
# separate column
get_municipalities(2002, 
                   geoid_year = 2018, 
                   geoid_ref_as_ref_column = TRUE) |> 
  head(n=5)
```

As an illustration, consider Lake Como, which was known as "South Belmar" before 2005, when it also had a different GEOID. We can see that it appears as a "new" municipality in the 2005 list:
```{r}
# Lake Como is the only 'new' municipality in 2005
dplyr::anti_join(
  get_municipalities(2005),
  get_municipalities(2004)
)
```
If we want to see the old name and GEOID, we can include the 2004 GEOID in the 2005 table and then join with the 2004 table:
```{r}
# Differences between 2005 and 2004
get_municipalities(2005, 
                   geoid_year = 2004, 
                   geoid_ref_as_ref_column = TRUE) |>
  dplyr::anti_join(get_municipalities(2004)) |>
  dplyr::left_join(get_municipalities(2004), 
                   by=c("GEOID_ref"="GEOID", "county"),
                   suffix=c("","_ref"))
```

#### Generating GEOID cross reference tables
Function `get_geoid_cross_references` will return a table of `GEOID` cross-references for a range of years to a specified reference year. Here we map all the GEOIDS from the years 2010-2020 to their 2005 GEOID.

```{r}
# Cross reference table, comparing GEOID for years 2010-2020 to 
# reference year 2005
get_geoid_cross_references(2005, 2010:2020) |>
  dplyr::arrange(GEOID_ref, year) |> 
  head(n=5)
```

#### Dealing with the Princetons
Princeton township and Princeton borough merged in 2013. Because the merged municipality retained Princeton borough's GEOID, Princeton township disappears in 2013. This is the only example of a disappearing municipality in the dataset.

The functions `get_municipality` and `get_geoid_cross_references` will return `NA` for a reference year GEOID after 2012 for Princeton township:
```{r}
# Princeton township existed in 2000 but not 2021
get_municipalities(2000, geoid_year = 2021) |>
  dplyr::filter(is.na(GEOID))
```

For convenience in dealing with this issue, this package includes the constants `PRINCETON_TWP_GEOID` and `PRINCETON_BORO_GEOID`.  
```{r}
c(njmunicipalities::PRINCETON_TWP_GEOID, njmunicipalities::PRINCETON_BORO_GEOID)
```
As an example, consider the municipal election data in [`njelections`](https://github.com/tor-gu/njelections). For comparisons across years, we may wish to combine the vote totals for the Princetons prior to 2013:
```{r}
library(njelections)
electoin_by_municipality_combined <-
  election_by_municipality |>
  dplyr::mutate(GEOID = dplyr::if_else(GEOID == PRINCETON_TWP_GEOID,
                                       PRINCETON_BORO_GEOID,
                                       GEOID)) |>
  dplyr::group_by(year, office, GEOID, party) |>
  dplyr::summarize(vote = sum(vote), .groups = "drop")
```

#### County list
For convenience, this package also includes a table of counties and
their GEOIDS. There have been no changes to counties from 2000 to 2021.
```{r}
# County list
counties |> head(n=5)
```

