---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
renv::use(
  "tor-gu/njmunicipalities"
)
```

# njmunicipalities

<!-- badges: start -->
[![R-CMD-check](https://github.com/tor-gu/njmunicipalities/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/tor-gu/njmunicipalities/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

This is a data package for R that contains every county
and municipality in New Jersey, from 2000 to 2021.

## Installation

You can install the development version of njmunicipalities from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("tor-gu/njmunicipalities")
```
## Examples
```{r}
library(njmunicipalities)

# Municipality list for 2007
get_municipalities(2007) |> head(n=5)

# Municipality list for 2002 with GEOIDs from 2018
get_municipalities(2002, geoid_year = 2018) |> head(n=5)

# Municipality list for 2002 with GEOIDs from 2018 added as
# seperate column
get_municipalities(2002, 
                   geoid_year = 2018, 
                   geoid_ref_as_ref_column = TRUE) |> 
  head(n=5)

# Cross reference table, comparing GEOID for years 2010-2020 to 
# reference year 2005
get_geoid_cross_references(2005, 2010:2020) |>
  dplyr::arrange(GEOID_ref, year) |> 
  head(n=5)

# County list
counties |> head(n=5)
```

## Changes in NJ municipalites, 2000-2021

Over the period 2000-2021, there have been several changes to
the list of municipalities in New Jersey:

* In 2005, South Belmar became Lake Como. The US Census assigned
a new GEOID to Lake Como.
* In 2007, Dover township in Ocean County became Toms River and was assigned a new GEOID.
* In 2008, Washington township in Mercer County became Robbinsville township and was assigned a new GEOID.
* In 2009, West Paterson became Woodland Park, and was assigned a new GEOID.
* In 2010, Caldwell borough was assigned a new GEOID from the the US Census, though there was no name change at this time.
* In 2013, Princeton borough and Princeton township merged. The merged municipality retained the Princeton borough GEOID, though the US Census started using the name 'Princeton' in place of 'Princeton borough' for the merged municipality.

